Postural changes mediate children's visual access to social information

### 2018-01-14 writeup-cogsci-copy/cogsci2018.Rmd
edits to paper

```{r, libraries}
library(knitr)
knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.crop = FALSE, 
                      fig.path='figs/', echo=FALSE, warning = FALSE, 
                      cache=TRUE, message = FALSE, sanitize = TRUE)
library(readr)
library(dplyr)
library(tidyr)
library(xtable)
library(lubridate)
library(langcog)
library(ggplot2)
library(lme4)
library(gridExtra)
```
# Introduction
Children are remarkably skilled language learners, connecting arbitrary labels (“cup”) with specific visual concepts at a rapid pace through the first two years of life. However, children do not learn words in a vacuum but rather in a rich social environment, where social cues provided by speakers (e.g., eye-gaze) provide strong scaffolding for this learning process. Further, children’s ability to effectively process these social cues may be a key factor in their early language development.  In a longitudinal study, children’s level of joint engagement with their mother was found to predict both their receptive and productive vocabularies (@carpenter1998).  More recently, 10 month-olds who follow an adult’s gaze in an experimental context have larger vocabularies at 18 months (@brooks2005) and throughout the second year of life (@brooks2008).

However, as children are learning their first words, their view of the world is also changing radically [@adolph2007]. Infant’s motor abilities improve dramatically near the end of the first year of life,  allowing them to locomote independently and to determine what they see. During spontaneous play in a laboratory playroom, toddlers are more likely to look at the floor while crawling than while walking (Franchak et al., 2011), when they have full visual access to their environment and the people in it (Kretch, Franchak, and Adolph, 2014).

@@

all_subs %>%
  left_join(inc_subs) %>%
  group_by(age_group) %>%
  summarise(n = sum(inc), 
            percent_included = mean(inc),
            age = mean(age.at.test, na.rm=TRUE), 
            video_length = mean(len.min, na.rm=TRUE), 
            female = sum(gender=="female", na.rm=TRUE)) %>%
  xtable %>%
  print(digits = 2, include.rownames=FALSE,table.placement = "H", type="latex", comment = F)
```
Our final sample consisted of 36 infants and children, with 8 participants in three age groups: 8 months (6 females), 12 months (7 females), and 16 months (6 females). Participants were recruited from the surrounding community via state birth records, had no documented disabilities, and were reported to hear at least 80\% English at home. Demographics and exclusion rates are given in Table \ref{tab:pop}. 
\begin{table}[H]
\centering
\begin{tabular}{rrrrrr}
  \hline
 Group & N & \% incl. & Mean age & Videos length (min) \\ 
  \hline
   8 mo. &   12 & 0.46 & 8.71 & 14.41 \\ 
   12 mo. &  12 & 0.40 & 12.62 & 13.48 \\ 
   16 mo. &  12 & 0.31 & 16.29 & 15.00\\ 
   \hline
\end{tabular}
\caption{\label{tab:pop} Demographics by age group.}
\end{table}
To obtain this final sample, we tested `r length(all_subs$reason)` children, excluding  `r sum(all_subs$reason!="included")` children for the following reasons:
`r sum(all_subs$reason=="HC technical error")` for technical issues related to the headcam,
`r sum(all_subs$reason=="would not wear camera")` for failing to wear the headcam,
`r sum(all_subs$reason=="less than 4 min HC footage")` for fewer than 4 minutes of headcam footage,
`r sum(all_subs$reason=="multiple people")` for having multiple adults present,
`r sum(all_subs$reason=="no CDI")` for missing CDI data,
`r sum(all_subs$reason=="no DV cam footage")` for missing scene camera footage,
`r sum(all_subs$reason=="infant crying")` for fussiness, and one excluded for sample symmetry. All inclusion decisions were made independent of the results of subsequent analyses.

@@

### Posture and Orientation Annotation

We created a set of custom annotations that described the child's physical posture (e.g. standing) and the orientation of the caregiver relative to the child (e.g. far away). The child's posture was categorized as being held/carried, prone (crawling or lying), sitting, or standing. The caregiver's orientation was characterized as being close to the child, farther from the child, and a global category of caregiver behind the child. For the first two annotationes (close/far from the child), the caregiver could either be to the the front or to the side of the child. All annontations were made using OpenSHAPA/Datavyu software  [@adolph2012], and times when the child was out of view of the tripod camera was marked as uncodable and was excluded from these annotations.

@@

The second algorithm was based on the work by (CITEXX) using multi-task cascaded convolutional neural networks (MTCNNs). The system was built using a novel cascaded CNN-based framework for joint detection and alignment, built to perform well in real-world environments where varying illuminations and occlusions are present. We used a Tensorflow implementation of this algorithm provided by (https://github.com/davidsandberg/facenet). Like Viola-Jones, this detector provided ifnormation about the presence of a face as well as it's size and position.

The third algorithm was a CNN-based pose detector that provided the locations of 18 body parts (ears, nose, wrists, etc) (CITEXX). The system uses a CNN for initial anatomical detection and subsequently applies part affinity fields (PAFs) for part association, producing a series of body part candidates. The candidates are then matched to a single individual and finally assembled into a pose. For the purposes of our investigation we only made use of the body parts relevant to the face (ears, eyes, nose). We operationalized face detections as ....., we operationalized hand detedctions as ..... we operationalized wrist detections as.....

### Detector evaluation

Evaluating face detectors (text here about how gold standard was selected)
Evaluating hand/wrist detectors (text here about how gold standard was selected)

@@

ggplot(detectionsByAge, aes(x = age.at.test, y = detections, color = detectorType, size=len.min)) +
    geom_jitter(width = .25) +
    geom_smooth() +
    scale_x_continuous(breaks=c(8,12,16)) +
    scale_size_continuous(range=c(.2,2),name = "Video Length (mins)") +
    ylab("Proportion Detections") +
    xlab("Age (months)") +
    scale_color_manual(values=c("#1F618D", "#7FB3D5", "#C70039", "#FF5733")) +
    facet_grid(~detectorType, scales = "free") +
    theme_bw()
```

>>>>

```{r, libraries}
library(knitr)
knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.crop = FALSE, 
                      fig.path='figs/', echo=FALSE, warning = FALSE, 
                      cache=TRUE, message = FALSE, sanitize = TRUE)
library(readr)
library(dplyr)
library(tidyr)
library(xtable)
library(lubridate)
library(langcog)
library(ggplot2)
library(lme4)
library(gridExtra)
```
# Introduction
Children are remarkably skilled language learners, connecting arbitrary labels (“cup”) with specific visual concepts at a rapid pace through the first two years of life. However, children do not learn words in a vacuum but in a rich social environment, where social cues provided by speakers (e.g., eye-gaze) provide strong scaffolding for this learning process. Indeed, children’s ability to effectively process these social cues may be a key factor in their early language development. For example,in a longitudinal study, children’s level of joint engagement with their mother was found to predict both their receptive and productive vocabularies (@carpenter1998).  More recently, 10 month-olds who follow an adult’s gaze in an experimental context have larger vocabularies at 18 months (@brooks2005) and throughout the second year of life (@brooks2008).

However, as children are learning their first words, their view of the world is also changing radically [@adolph2007]. Infant’s motor abilities improve dramatically near the end of the first year of life,  allowing them to locomote independently and to determine what they see. These motor changes have drastic consequences for what children see: crawling and walking infants have radically different views of the world. During spontaneous play in a laboratory playroom, toddlers are more likely to look at the floor while crawling than while walking (Franchak et al., 2011); walking vs. crawling infants tend to have full visual access to their environment and the people in it (Kretch, Franchak, and Adolph, 2014).

@@

all_subs %>%
  left_join(inc_subs) %>%
  group_by(age_group) %>%
  summarise(n = sum(inc), 
            percent_included = mean(inc),
            age = mean(age.at.test, na.rm=TRUE), 
            video_length = mean(len.min, na.rm=TRUE), 
            female = sum(gender=="female", na.rm=TRUE)) %>%
  xtable %>%
  print(digits = 2, include.rownames=FALSE,table.placement = "H", type="latex", comment = F)
```

Our final sample consisted of 36 infants and children, with 8 participants in three age groups: 8 months (6 females), 12 months (7 females), and 16 months (6 females). Participants were recruited from the surrounding community via state birth records, had no documented disabilities, and were reported to hear at least 80\% English at home. Demographics and exclusion rates are given in Table \ref{tab:pop}. 
\begin{table}[H]
\centering
\begin{tabular}{rrrrrr}
  \hline
 Group & N & \% incl. & Mean age & Videos length (min) \\ 
  \hline
   8 mo. &   12 & 0.46 & 8.71 & 14.41 \\ 
   12 mo. &  12 & 0.40 & 12.62 & 13.48 \\ 
   16 mo. &  12 & 0.31 & 16.29 & 15.00\\ 
   \hline
\end{tabular}
\caption{\label{tab:pop} Demographics by age group.}
\end{table}
To obtain this final sample, we tested `r length(all_subs$reason)` children, excluding  `r sum(all_subs$reason!="included")` children for the following reasons:
`r sum(all_subs$reason=="HC technical error")` for technical issues related to the headcam,
`r sum(all_subs$reason=="would not wear camera")` for failing to wear the headcam,
`r sum(all_subs$reason=="less than 4 min HC footage")` for fewer than 4 minutes of headcam footage,
`r sum(all_subs$reason=="multiple people")` for having multiple adults present,
`r sum(all_subs$reason=="no CDI")` for missing CDI data,
`r sum(all_subs$reason=="no DV cam footage")` for missing scene camera footage,
`r sum(all_subs$reason=="infant crying")` for fussiness, and one excluded for sample symmetry. All inclusion decisions were made independent of the results of subsequent analyses.

@@

### Posture and Orientation Annotation
We created a set of custom annotations that described the child's physical posture (e.g. standing) and the orientation of the caregiver relative to the child (e.g. far away). The child's posture was categorized as being held/carried, prone (crawling or lying), sitting, or standing. The caregiver's orientation was characterized as being close to the child, farther from the child, and a global category of caregiver behind the child. For the first two annotationes (close/far from the child), the caregiver could either be to the the front or to the side of the child. All annontations were made using OpenSHAPA/Datavyu software  [@adolph2012], and times when the child was out of view of the tripod camera was marked as uncodable and was excluded from these annotations.

@@

The second algorithm was based on the work by (CITEXX) using multi-task cascaded convolutional neural networks (MTCNNs). The system was built using a novel cascaded CNN-based framework for joint detection and alignment, built to perform well in real-world environments where varying illuminations and occlusions are present. We used a Tensorflow implementation of this algorithm provided by (https://github.com/davidsandberg/facenet). Like Viola-Jones, this detector provided information about the presence of a face as well as it's size and position.

The third algorithm was a CNN-based pose detector that provided the locations of 18 body parts (ears, nose, wrists, etc) (CITEXX). The system uses a CNN for initial anatomical detection and subsequently applies part affinity fields (PAFs) for part association, producing a series of body part candidates. The candidates are then matched to a single individual and finally assembled into a pose. For the purposes of our investigation we only made use of the body parts relevant to the face (ears, eyes, nose). We operationalized face detections as any frames containing a nose and either the left or right eye; we operationalized hand detections as any frames containing any of the 21 points detected on either the left or the right hand, and we operationalized wrist detections any frames containing either the left or the right wrist.

### Detector evaluation
In order to evaluate the performance of these detectors, we constructed a "gold set" of frames

@@

ggplot(detectionsByAge, aes(x = age.at.test, y = detections, color = detectorType, size=len.min)) +
    geom_jitter(width = .25) +
    geom_smooth(size=1) +
    scale_x_continuous(breaks=c(8,12,16)) +
    scale_size_continuous(range=c(.01,2),name = "Video Length (mins)") +
    ylab("Proportion Detections") +
    xlab("Age (months)") +
    scale_color_manual(values=c("#1F618D", "#7FB3D5", "#C70039", "#FF5733")) +
    facet_grid(~detectorType, scales = "free_y") 
    # theme_bw()
```
